Apparent preferences for cognitive effort fade when multiple forms of effort

### 2020-11-05 Cost_manuscript.Rmd
separated stability and modeling sections
new consistency and stability exp 1

  We next tested the hypothesis that foragers would display consistent and stable choices. Since participants had complete a priori knowledge of the statistics of each environment, and were unaware of any other cost groups that could modulate the subjective cost of their assigned demand, we expected them to 1) show confidence in their choices by not quitting mid-handling time; and 2) adopt the same strategy while re-experiencing each environment. We investigated the first prediction by plotting a survival curve of trial quits across all participants (Figure 5, top-left), with black crosses signaling either completed trials (at 2 s, 10 s, and 14 s) or forced travels. This qualitative analysis shows that most decisions to quit happened within the first second into the handling time, and participants rarely quit during the handling time. Physical and easy groups show slightly lagged responses because the experiment allowed a one-second grace period for participants to begin gripping (marked by the dashed line). This resulted in some individuals choosing to wait for that second to indicate an offer rejection (the censor at 4s in the physical group was a completion during a 2 s handling trial that was incorrectly recorded as 4 s by the experimental code). Plotting the quitting time distributions across blocks for the first handling second (ECDFs on Figure 5, top-right) shows that participants in the wait and cognitive groups responded faster as the experiment progressed. These response time observations suggest clarity in participant preferences.

@@

  We tested the second prediction (i.e. strategy stability) by comparing acceptances rates throughout the experiment. The scatter plot on Figure 5 (bottom-left) shows the overall proportion accepted by each participant before and after the mid-point break (colored dashed lines indicate the mean for that group). Paired permutations showed no acceptance rate differences for wait and easy groups (*p* > 0.05), but we found significantly lower acceptances for physical (*p* `r report_p(prepost_btw$Perms[1])`, Cohen's D = `r round(prepost_btw$CohenD[1], digits = 2)`) and cognitive (*p* `r report_p(prepost_btw$Perms[2])`, Cohen's D = `r round(prepost_btw$CohenD[2], digits = 2)`) groups. The decline in acceptances appeared to be gradual for the cognitive group, but not for the physical group (Figure 5, bottom-right). An explanation for these results is that these conditions became costlier due to fatigue, thus providing evidence that our cognitive and physical manipulations were effortful.

  It is possible that this reduction in acceptances affected the relative differences among conditions. We addressed this possibility by examining pre-post acceptance rates for each handling time and reward combination. Figure 6 (left) shows the proportion of acceptances for each offer combination, with each half depicted on each column. This collage suggests that changes in choice behavior for effortful groups were mostly driven by the longest handling time, in in line with reports where value is discounted as the length of the required effort increases (Treadway et al, 2009). Regardless, the relative differences in acceptances among cost groups remain mostly untouched. We tested this observation by repeating the winning mixed-effects model from the previous section (i.e. all main effects) on each experimental half separately. We replicated the matrix of coefficients from the previous section (Figure 6, right), with each half mirrored along the diagonal (i.e. the top triangle shows the coefficients for the second half of the experiment). Together, these results suggest that decision makers had a strong representation of the environment by showing that 1) participants were confident in their choices and strategies; 2) cognitive and physical costs produced the expected fatigue from exerting effort; and 3) despite the resulting reduction of acceptances, the relative cost of each condition was consistent throughout the experiment.

@@

  Together, these findings suggest that 1) foraging paradigms can provide a reference point to evaluate how different demands modulate value; 2) a simple OC model can approximate people's behavior in such single-option, take it or leave it choices; and 3) when foraging in environments of equivalent richness, cognifitvely effortful demands can increase the attractiveness of offers rather than discount them (unlike other forms of cost). In the next experiment we explored whether this apparent value was inherent or malleable.

@@

## Discussion

>>>>

  Next, we tested the hypothesis that foragers would display consistent and stable choices. Participants were deemed consistent if they did not quit during the handling time. To examine this, we plotted survival curves of trial quits across all participants (Figure 5, top-left), with black crosses signaling either completed trials (at 2 s, 10 s, and 14 s) or forced travels. These curves show that most decisions to quit happened within the first second into the handling time, and participants rarely quit during the handling time. Physical and easy groups show slightly lagged responses because the experiment allowed a one-second grace period for participants to begin gripping (marked by the dashed line). This resulted in some individuals choosing to wait for that second to indicate an offer rejection (the censor at 4s in the physical group was a completion during a 2 s handling trial that was incorrectly recorded as 4 s by the experimental code). Plotting the quitting time distributions across blocks for the first handling second (ECDFs on Figure 5, top-right) shows that participants in the wait and cognitive groups responded faster as the experiment progressed. These response time observations suggest clarity in participant preferences.

@@

  Stability was defined as a persistence in acceptance patterns throughout the experimental session. The scatter plot on Figure 5 (bottom-left) shows the overall proportion accepted by each participant before and after the mid-point break. Paired permutations showed no differences in acceptances for wait and easy groups (*p* > 0.05), but we found significantly lower acceptances for physical (*p* `r report_p(prepost_btw$Perms[1])`, Cohen's D = `r round(prepost_btw$CohenD[1], digits = 2)`) and cognitive (*p* `r report_p(prepost_btw$Perms[2])`, Cohen's D = `r round(prepost_btw$CohenD[2], digits = 2)`) groups. Figure 6 (left) shows that changes in choice behavior for effortful groups were mostly driven by the longest handling time, in in line with reports where value is discounted as the length of the required effort increases (Treadway et al, 2009). We replicated the mixed-effects logistic regression from the previous section separately on each experimental half, which confirmed that differences among demands were stable (Figure 6, right). Together, the consistency and stability displayed by participants suggest that decision makers had a clear representation of the environment.

@@

@@

### **Computational modeling of foraging behavior**

  Together, these findings suggest that 1) foraging paradigms can provide a reference point to evaluate how different demands modulate value; 2) a simple OC model can approximate people's behavior in such single-option, take it or leave it choices; and 3) when foraging in environments of equivalent richness, cognifitvely effortful demands can increase the attractiveness of offers rather than discount them (unlike other forms of cost). In the next experiment we explored whether this apparent value was inherent or malleable.
## Discussion
